A new method to estimate dark matter halo concentrations

### 2015-10-19 paper/paper.tex
intro

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc}\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc}\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
%SAMPLES
\newcommand{\GHBDM}{\texttt{GH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\GHFOF}{\texttt{GH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IHBDM}{\texttt{IH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IHFOF}{\texttt{IH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\PBDM}{\texttt{P}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\PFOF}{\texttt{P}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IPBDM}{\texttt{IP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IPFOF}{\texttt{IP}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\RIPBDM}{\texttt{RIP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\RIPFOF}{\texttt{RIP}$_{\mbox{\tiny{FOF}}}$ }

@@

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{A new method to estimate dark matter halo concentrations}
\author[C. Poveda, J.E. Forero-Romero & J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$,
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$,
  J. C. Mu\~noz-Cuartas \thanks{xxxx@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

@@

%*************************************************************************
\section{Introduction}
\label{sec:introduction}
%*************************************************************************

\citep{NFW,Taylor2001}

@@

\end{document}
The tracer that we use to define the voids is the fractional
anisotropy (FA) computed from the set of eigenvalues of the tensor
under consideration. 
The FA was initially introduced by \citet{Basser95} to quantify the
anisotropy degree of the diffusivity of water molecules through
cerebral tissue in nuclear magnetic resonance imaging, thereby allowing to 
detect structures that restrict the otherwise isotropic Brownian movement 
of water molecules. Taking into account that this process is identical to 
the velocity shear in the large-scale matter distribution, 
\citet{Libeskind13} introduced the concept of the FA in the context of 
Cosmic Web classification schemes.

>>>>

\newcommand{\LCDM}{$\Lambda$CDM~}
\newcommand{\beq}{\begin{eqnarray}}  
\newcommand{\eeq}{\end{eqnarray}}  
\newcommand{\zz}{$z\sim 3$} 
\newcommand{\apj}{ApJ}  
\newcommand{\jcap}{JCAP}  
\newcommand{\apjs}{ApJS}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha}\else{Ly$\alpha$}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc}\fi}}  
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc}\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc}\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}  
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
%SAMPLES
\newcommand{\GHBDM}{\texttt{GH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\GHFOF}{\texttt{GH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IHBDM}{\texttt{IH}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IHFOF}{\texttt{IH}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\PBDM}{\texttt{P}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\PFOF}{\texttt{P}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\IPBDM}{\texttt{IP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\IPFOF}{\texttt{IP}$_{\mbox{\tiny{FOF}}}$ }
\newcommand{\RIPBDM}{\texttt{RIP}$_{\mbox{\tiny{BDM}}}$ }
\newcommand{\RIPFOF}{\texttt{RIP}$_{\mbox{\tiny{FOF}}}$ }

@@

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title{A new method to estimate dark matter halo concentrations}
\author[C. Poveda, J.E. Forero-Romero, J.C. Mu\~noz-Cuartas]{
\parbox[t]{\textwidth}{\raggedright 
  C. Poveda \thanks{cn.poveda542@uniandes.edu.co}$^{1}$
  J. E. Forero-Romero \thanks{je.forero@uniandes.edu.co}$^{1}$
  J. C. Mu\~noz-Cuartas \thanks{xxxx@udea.edu.co}$^{2}$
}
\vspace*{6pt}\\
$^1$Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
$^2$Instituto de F\'{\i}sica - FCEN, Universidad de Antioquia, Calle
67 No. 53-108, Medell\'{\i}n, Colombia
}

@@

%*************************************************************************
\section{Introduction}
\label{sec:introduction}
%*************************************************************************
In the concordance cosmological paradigm the matter content of the
Universe is dominated by dark matter, a collisionless fluid shaped by
gravitational interactions.
In the last three decades simulations of dark matter dominated
universes have provided valuable insights into the large scale
structure formation process, showing a remarkable success in the
comparison between theoretical results and observations of the
galaxy distribution obtained from surveys.
\citep{Springel2005,2011ApJ...740..102K}.

On galactic scales the most striking result is that dark matter
overdensities closely follow a universal density profile. 
In a first approximation this profile is spherically symmetric and its
density only dependens on the radial coordinate.
The universality of this profiles seems to be independent of the
cosmological parameters and is self-similar for different spatial
scales after an adequate re-scaling is applied. 
\citep{NFW,Taylor2001}

One the most popular parameterization for a dark matter halo radial
density distribution is the Navarro-Frenk-White (NFW) profile
\citep{NFW}.
This profile is a double power law in radius, where the transition break
happens at the so-called scale radius, $r_s$.
The ratio between the scale radius and the virial radius $R_v$, which
defines a natural scale for the halo, is known as the concentration
$c=R_v/r_s$. 
Simulations also show that the concentration is a strong function of
halo mass and redshift.

However, high resolution simulations of Milky Way sized dark matter halos
\citep{Navarro2010} show that the universality property is not
perfect and that a better fitting parameterization to the radial
density profile is provided by the Einasto profile
\citep{Einasto1965}.
Nevertheless, the NFW density profile and its concentration have
become a standard metric to describe the structure of dark matter
halos.

Observationally, the relationship between halo mass and concentration
could provide a potential test of the Cold Dark Matter (CDM) paradigm
on galactic scales. 
For this motivation a great deal of effort has been
invested in calibrating this relationship with simulations
\citep{Neto2007,Duffy2008,Munoz2011,Prada2012,Ludlow2014} and 
finding the best possible way to constraint it with observations
\citep{Buote2007,Comerford2007,Mandelbaum2008,Giocoli2014,Foex2014,Shan2015}.

In N-body simulations there are two main methods to estimate the
concentration parameter of a dark matter halo in a N-body simulation.
The first method takes the particles composing the halos
and bins them in logarithmic radii to estimate the density in each
bin, then it proceeds to fit the density as a function of the radius
to the NFW profile.
A second method uses an analytic property of the
NFW profile that relates the maximum of the ratio of the circular velocity to
the virial velocity.
The concentration can be then found as the root of a algebraic 
equation dependent on this maximum value.

The first method is straightforward to apply but presents two
disadvantages.
First, it requires a large number of particles in
order to have a proper density estimate in each bin.
This makes the method robust only for halos with at least  $10^3$
particles.
The second problem is that there is not a way to estimate the optimal
radial bin size, different choices produce different results for the
concentration.

The second method solves the two problems mentioned above.
It works with low particles numbers and does not involve data
binning.
However, it effectively takes into account only a single data point
and discards the behaviour of the ratio $V_{\rm circ}/V_{\rm vir}$ below and
above its maxima.
Additionally, small fluctuations on the value of this maximum can
yield large perturbations on the estimated concentration parameter.

In this paper we propose a new method to estimate the dark matter halo
concentration in halos from N-body simulations.
Our method builds the cumulative mass profile from the particle
to find the best possible concentration value fitting this
distribution. 
This proposal has two advantages with respect to the two methods mentioned
above. 
It does not involve any data binning and does not throw away data points.

@@

The tracer that we use to define the voids is the fractional
anisotropy (FA) computed from the set of eigenvalues of the tensor
under consideration. 
The FA was initially introduced by \citet{Basser95} to quantify the
anisotropy degree of the diffusivity of water molecules through
cerebral tissue in nuclear magnetic resonance imaging, thereby allowing to 
detect structures that restrict the otherwise isotropic Brownian movement 
of water molecules. Taking into account that this process is identical to 
the velocity shear in the large-scale matter distribution, 
\citet{Libeskind13} introduced the concept of the FA in the context of 
Cosmic Web classification schemes.
